\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-02 paper/main.tex
Updates in Results 4

Ciliary motion is integral to physiological CSF flow,
and absence of cilia has been associated with defective
brain development and various pathologies~\cite{Eichele2020Cilia-drivenVentricle,
Olstad2019CiliaryDevelopment, Faubel2016Cilia-basedVentricles,
Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Sawamoto2006NewBrain, Yoshiba2014RolesSymmetry,
Hirokawa2006NodalAsymmetry}.
As observed both experimentally and computationally,
absence of motile cilia alters CSF flow patterns. 
We next investigate the implications of these changes on intraventricular
transport.

We experimentally measured the movement of proteins by monitoring
photoconverted Dendra2 signal in control zebrafish and 
motile-cilia mutant \emph{schmalhans} (\emph{smh}) zebrafish.
In ROIs 2 and 4, which quantify Dendra2 distribution
downstream of the photoconversion site, we observed a slower increase
of fluorescence intensity (\Crefpanels{fig:fig4}{a}). The mean times-to-threshold
in the \emph{smh} mutants were $45\%$ (ROI 2) and $25\%$ (ROI 4) higher
compared to controls (\Crefpanels{fig:fig4}{b}). We found no significant differences
between \emph{smh} mutants and controls
in the Dendra2 distribution in ROIs 3, 5, and 6.

The \emph{smh} mutants are mimicked in the computational model by
considering transport of Dendra2 with CSF flow driven by cardiac
pressure pulsations only, representing the absence of motile
cilia. Comparing the in silico mutants with transport predictions in
the presence of both cardiac pulsations and motile cilia, we observe
noticeably delayed dynamics in ROIs 2, 4 and 5, but little change for
ROIs 3 and 6 (\Crefpanels{fig:fig4}{c}).

Compared to experimental data, the difference between the control and
mutant scenarios was even greater for the simulations, with the
times-to-threshold in ROI 2 and 4 being 81\% and 28\% higher in the
absence of cilia-mediated flow. Moreover, the ROI 5 $\cbar$ profiles
for the two configurations in our computational model deviate,
which is not the case in the experiments. Altogether,
our findings identify that, for smaller proteins such as Dendra2,
diffusion plays an important role for
solute movement to regions upstream in the cilia-mediated flow fields or to
distant ventricles, while cilia affect the downstream mean concentrations.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $\Delta F(t)$
    in each region of interest (ROI) for control (black lines, $n=8$)
    and \emph{smh} mutant (red lines, $n=6$) fish. 
    Bold lines are mean values of the cohorts and shaded regions cover one standard devation errors
    in the means.
    \textbf{b.} Times when $\Delta F(t)$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for control (gray dots) and
    \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets. The $p$-values
    were calculated using a Mann-Whitney U test with a confidence level of 95\%, with 
    the null hypothesis that the two distributions of control and mutant data are 
    from the same population.
    \textbf{c.} Mean concentration $\cbar(t)$ in each ROI,
    simulated using the baseline model (black lines) and the cardiac-induced/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have
    different limits in each plot.}
    \label{fig:fig4}
\end{figure}

>>>>

Ciliary motion is integral to physiological CSF flow, and absence of
cilia has been associated with defective brain development and various
pathologies~\cite{Eichele2020Cilia-drivenVentricle,
  Olstad2019CiliaryDevelopment, Faubel2016Cilia-basedVentricles,
  Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
  Sawamoto2006NewBrain, Yoshiba2014RolesSymmetry,
  Hirokawa2006NodalAsymmetry}.  As observed both experimentally and
computationally, absence of motile cilia alters CSF flow patterns.  We
next investigate the implications of these changes on intraventricular
transport. We experimentally measured the movement of proteins by
monitoring photoconverted Dendra2 signal in control zebrafish and
motile-cilia mutant \emph{schmalhans} (\emph{smh}) zebrafish.  In ROIs
2 and 4, which quantify Dendra2 distribution downstream of the
photoconversion site, we observed a slower increase of fluorescence
intensity (\Crefpanels{fig:fig4}{a}): the mean times-to-threshold in
the \emph{smh} mutants were $45\%$ (ROI 2) and $25\%$ (ROI 4) higher
compared to controls (\Crefpanels{fig:fig4}{b}). We found no
significant differences between \emph{smh} mutants and controls in the
Dendra2 distribution in ROIs 3, 5, and 6.

In the computational model, we represent the \emph{smh} mutants by
considering transport of Dendra2 with CSF flow driven by cardiac
pressure pulsations only, representing the absence of motile
cilia. Comparing the in silico mutants with transport predictions in
the presence of both cardiac pulsations and motile cilia, we observe
noticeably delayed dynamics in ROIs 2, 4 and 5, but little change for
ROIs 3 and 6 (\Crefpanels{fig:fig4}{c}). Compared to experimental
data, the control and mutant scenarios differed more for the
simulations, with the times-to-threshold in ROI 2 and 4 being 81\% and
28\% higher in the absence of cilia-mediated flow. Moreover, the
predicted mean concentrations in ROI 5 deviate between the two
scenarios, which is not the case in the experiments. Altogether, our
findings identify that, for smaller proteins such as Dendra2,
diffusion plays an important role for solute movement to regions
upstream in the cilia-mediated flow fields or to distant ventricles,
while the cilia affect the downstream mean concentrations.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $\Delta F(t)$
    in each region of interest (ROI) for control (black lines, $n=8$)
    and \emph{smh} mutant (red lines, $n=6$) fish. 
    Bold lines are mean values of the cohorts and shaded regions cover one standard devation errors
    in the means.
    \textbf{b.} Times when $\Delta F(t)$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for control (gray dots) and
    \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets. The $p$-values
    were calculated using a Mann-Whitney U test with a confidence level of 95\%, with 
    the null hypothesis that the two distributions of control and mutant data are 
    from the same population.
    \textbf{c.} Mean concentration $\cbar(t)$ in each ROI,
    simulated using the baseline model (black lines) and the cardiac-induced/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have
    different limits in each plot.}
    \label{fig:fig4}
\end{figure}
